---
title: "Preterm Exercise"
author: "Carole Voulgaris"
date: '2022-08-16'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
The purpose of this exercise is to give you some practice with the RStudio
interface and a few common tidyverse functions.

## Required packages
This exercise uses the following packages:

```{r, message=FALSE, results='hide'}
library(tidyverse)
library(sf)
library(here)
```

## Load data from a file

Start an RStudio project.

Download the csv file called "FY2022_cambridge_asssessor.csv" from Canvas 
(Pre-Term Workshop > R exercise) and save it to the project directory associated
with the RStudio project you just created.

Create a new R script in your RStudio project. In the first few lines of the 
script, load the required packages. Somewhere below that, add the following 
lines to your script:

```{r, results='hide', message=FALSE}
parcel_data <- here("FY2022_cambridge_assessor.csv") %>%
  read_csv()
```

The `here()` function will create a path to csv file in your project directory.
The pipe (`%>%`) at the end of that line will take that path and enter it as 
the first argument of the `read_csv()` function on the next line. The `read_csv()`
function is aptly named. It reads a csv file and saves the contents to a tibble
you've named at the start of this piped set of functions (`parcel_data`)

## Select columns

You should see the `parcel_data` tibble listed on your environment
tab now. It has 29,542 rows (observations) and 63 columns (variables).
Not all of these variables are interesting or interpretable for us. Let's
try selecting just the ones we want to keep.

```{r}
owners_values <- parcel_data %>%
  select(gisid, landarea, buildingvalue, landvalue, owner_name)
```

Now you should have a new tibble in your environment called `owners_values`.
It has all the same rows as parcel_data, but only includes the five 
variables you selected using the `select` function.

## Filter rows

Maybe we only want to analyze the properties owned by Harvard.
We can filter for these. 

In the code below, the `filter()` function, will keep only the 
rows that meet the a specified condition. The double equal sign (`==`)
means you want to compare the two values (rather than setting one 
equal to the other, which is what a single equal sign would suggest).

```{r}
Harvard_parcels <- owners_values %>%
  filter(owner_name == "PRESIDENT & FELLOWS OF HARVARD COLLEGE")
```

That leaves 372 Harvard-owned properties. Glancing through the dataset, some
of these appear to be duplicates. We can remove those with the `filter()` 
function as well.

```{r}
Harvard_unique <- Harvard_parcels %>%
  filter(!duplicated(.))
```

A couple of translations in the above code:

* `!` means "not"
* `.` references the object that's getting piped in from the 
previous line

## Calculate a new variable


## Join two datasets


## Plot some data


## Save a file


